What's hot and what's not:\\Windowed developer topic analysis

### 2009-06-23 lda-paper.tex
some ric changes to the LDA paper

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  \hla{
    Latent Semantic Indexing (LSI) can be used to analyze commit log
  }
  comments over the entire lifetime of a project. Previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA  to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose 
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

\hla{ The questions that we wish to address deal with the past and
  topics of development that occurred in the past.  What happened in
  the previous iteration of a project's development?  What were the
  developers working on? What was the developer in the next cubicle
  working on? Which common topics and issues were dealt with in the
  current release of the software? What were the topics of the
  previous release? Given the requirements agreed to at the start of
  the iteration did our developers work on them?  What else did they
  work on?  }

\hla{
We want to know about the topics of development that programmers
encountered while developing a software project. We want to know given
a certain period of time, what were the popular topics being worked on
during that period. We also want to know if any of those topics
occurred again, whether before or after our current period.  These
topics could help stakeholders determine what their coworkers were
working on, what were common issues, what requirements were being
worked on, etc. In order to extract and evaluate these topics we plan
to rely on topic analysis.
}

Topic analysis extracts groups of words from documents (commit log
comments) that are word distributions, topics, that are combined to
characterize clusters of documents. These extracted topics often
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
\hla{
message, to be related to multiple topics. This maps well to commits
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS). 
}

\hla{
Usually topic analysis is applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
}
where as we suggest that more topics are probably local, if a
programmer is working on a story card, they might never deal with that
story card ever again. Thus that topic might be quite relevant to a
small window of time but never relevant ever again.  By applying topic
analysis locally to an iteration, both the developers and managers
might be able to see what were the topics of development during that
iteration, rather than the whole project.

\hlb{
We want to see topics that are unique to a
certain time period, displayed across the time axis.
Figure \ref{fig:lda} illustrates the kind of visualization we want to
automatically generate. } Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a single word drawn from its
word distribution.  
\hdb{
Given a visualization, such as Figure \ref{fig:lda}, topic
analysis can aid in partitioning a project's time-line into periods of
developer focus.
}
By breaking apart an iteration into sets of topics
and trends (topics that recur), we may be able to recognize the
underlying software development process and maintenance tasks from these
commits.

>>>>

  As development on a software project progresses, developers shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  \hla{
    Latent Semantic Indexing (LSI) can be used to analyze commit log
  }
  comments over the entire lifetime of a project. Previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this paper, we use LDA  to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose 
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a project's lifetime because many topics are quite local.

@@

\hla{ The questions that we wish to address deal with the past and
  topics development that occurred in the past.  What happened in
  the previous iteration of a project's development?  What were the
  developers working on? What was the developer in the next cubicle
  working on? Which common topics and issues were dealt with in the
  current release of the software? What were the topics of the
  previous release? Given the requirements agreed to at the start of
  the iteration did our developers work on them?  What else did they
  work on?  }

\hla{
We want to know about the topics of that programmers
encountered while developing a software project. We want to know given
a certain period of time, what were the popular topics being worked on
during that period. We also want to know if any of those topics
occurred again, whether before or after our current period.  These
topics could help stakeholders determine what their coworkers were
working on, what were common issues, what requirements were being
worked on, etc. In order to extract and evaluate these topics we plan
to rely on topic analysis.
}

\hdb{
Topic analysis extracts independant word distributions (topics) from
documents (commit log comments).
Ideally these extracted topics }
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
\hlb{
message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
}
\hla{
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS). 
}

\hla{
In the past, topic analysis has been applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
}
\hlb{
where-as this paper suggests that most topics are local. For example,  
programmer may be working on a story card,} they might never deal with that
story card ever again. Thus that topic might be quite relevant to a
small window of time but never relevant ever again.  By applying topic
analysis locally to a window or an iteration, both the developers and managers
might be able to see what were the topics of development during that
iteration, rather than the whole project.

\hlb{
We want to see topics that are unique to a
certain time period, displayed across the time axis.
Figure \ref{fig:lda} illustrates the kind of visualization we want to
automatically generate. } Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a word drawn from its
word distribution.  
\hdb{
Given a visualization, such as Figure \ref{fig:lda}, topic
analysis can aid in partitioning a project's time-line into periods of
developer focus.
}
By breaking apart an iteration into sets of topics
and trends (topics that recur), we may be able to recognize the
underlying software development process and maintenance tasks from these
commits.
